---
title: "Module1_0"
author: "AdrienM"
date: "2026-05-21"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Workshop 1 - Introduction to R and RStudio

```{r}
source(here::here("Workshop 1", "workshop1.R"))
```

# Worshop 2 - Data Visualization with ggplot2

#New chunk shortcuts 

Comd + opt I 
Crtl + Alt I 


```{r}
library(tidyverse)
```

```{r}
mpg
```

#first ggplot
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = class))
```
#change point color by class
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```
#change point shape by class
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

```{r}
library(ggplot2)

ggplot(data = mpg) +
  geom_point(
    mapping = aes(
      x = displ,
      y = hwy,
      colour = displ < 5
    )
  )
```
#facet and panel plots
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```
#Fitting simple lines
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) +
#  geom_point(mapping = aes(x = displ, y = hwy)) # points horrible
  geom_smooth(mapping = aes(x = displ, y = hwy)) # try smooth line
```

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```

```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE,
  )
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth()
```


```{r}
library(dplyr)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth(
    data = filter(mpg, class == "subcompact"),
    se = FALSE
  )
```

```{r}
 ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```
#Transformations and Stats
#Plotting statistics

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

```{r}
ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
```

#overriding defaults in ggplot2
```{r}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)
demo

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))
#> Warning: `stat(prop)` was deprecated in ggplot2 3.4.0.
```

#Plotting statistical details

```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

#Aestestic adjustements

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
#To alter transparency (alpha)
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 1/5, position = "identity")

#To color the bar outlines with no fill color
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) +
  geom_bar(fill = NA, position = "identity")
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

#The layered grammar of graphics

```{r}
ggplot(data = mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = class),
    stat = "identity",
    position = "identity"
  )
```

<!-- ggplot(data = <DATA>) +  -->
<!--   <GEOM_FUNCTION>( -->
<!--      mapping = aes(<MAPPINGS>), -->
<!--      stat = <STAT>,  -->
<!--      position = <POSITION> -->
<!--   ) + -->
<!--   <FACET_FUNCTION> -->

# Workshop 3 - R studio to github 

```{r}
library(usethis)
```

```{r}
use_git_config(user.name = "MESSEAUD13", user.mail = "adriencharleslouis.messeaud@my.jcu.edu.au")
```

# Workshop 4 

```{r}
# Load tidyverse (if not already loaded)
library(tidyverse)
```

```{r}
# Sample coral data
coral_monitoring <- tibble(
  site = c("Site_A", "Site_B", "Site_C"),
  depth_m = c(12, 35, 8)
)

# Classify depth using if_else
coral_monitoring <- coral_monitoring %>% 
  mutate(zone = if_else(depth_m > 30, "Deep Reef", "Shallow Reef"))
```

```{r}
coral_monitoring <- coral_monitoring %>% 
  mutate(reef_category = case_when(
    depth_m < 10  ~ "Lagoon / Flats",
    depth_m <= 30 ~ "Crest / Slope",
    depth_m > 30  ~ "Mesophotic / Deep",
    TRUE          ~ "Unclassified" # Catch-all remainder
  ))
```

```{r}
# Classic loop anatomy
for (year in 2020:2024) {
  print(paste("Processing climate data for year:", year))
}
```

```{r}
# Loop across index sequences
transect_lengths <- c(50, 100, 25, 75)

for (i in seq_along(transect_lengths)) {
  print(paste("Transect number", i, "measures", transect_lengths[i], "metres."))
}
```

```{r}
site_areas <- c(144, 400, 625)
results <- numeric(length(site_areas)) # Must build an empty container first

for(i in seq_along(site_areas)) {
  results[i] <- sqrt(site_areas[i])}
```

```{r}
library(purrr)
# Map directly and define your expected output type explicitly
mapped_results <- map_dbl(site_areas, sqrt)
```

```{r}
# Iterate a summary function over split lists of data
iris %>% 
  split(.$Species) %>% 
  map(~summary(.x))
```

```{r}
# Function Definition
calculate_coral_mortality <- function(initial_count, surviving_count) {
  
  # Logic safety switch using our conditional tools!
  if (initial_count <= 0) {
    stop("Initial count must be greater than zero.")
  }
  
  mortality_rate <- (initial_count - surviving_count) / initial_count
  return(mortality_rate)
}
```

# Utilizing your custom function
```{r}
calculate_coral_mortality(initial_count = 120, surviving_count = 84)
```

```{r}
convert_temp_c_to_f <- function(celsius) {
  if (celsius < -273.15) {
    stop("Temperature cannot be below absolute zero (-273.15°C).")
  }
  fahrenheit <- (celsius * 9/5) + 32
  return(fahrenheit)
}
```

```{r}
# Valid test
convert_temp_c_to_f(25)
```

```{r}
library(tidyverse)
library(palmerpenguins)

# Load the penguins dataset
data(penguins)

# Create custom color palettes
species_colors <- c(
  "Adelie" = "#FF6B6B",
  "Chinstrap" = "#4ECDC4",
  "Gentoo" = "#45B7D1"
)

# Generate boxplot comparing flipper_length_mm across species, faceted by island
penguins %>%
  drop_na(flipper_length_mm) %>%
  ggplot(aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_boxplot(alpha = 0.8, outlier.size = 2, outlier.alpha = 0.7) +
  facet_wrap(~island, ncol = 3) +
  scale_fill_manual(values = species_colors) +
  labs(
    title = "Flipper Length Distribution by Species and Island",
    subtitle = "Palmer Penguins Dataset",
    x = "Species",
    y = "Flipper Length (mm)",
    fill = "Species"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "gray60"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 11, face = "bold"),
    panel.grid.major.y = element_line(color = "gray90"),
    legend.position = "bottom"
  )

# Optional: Save the plot
ggsave("flipper_length_boxplot.png", width = 12, height = 6, dpi = 300)

```

```{r}
library(tidyverse)
library(palmerpenguins)
```

```{r}
# Load the penguins dataset
data(penguins)

# Create custom color palettes
species_colors <- c(
  "Adelie" = "#FF6B6B",
  "Chinstrap" = "#4ECDC4",
  "Gentoo" = "#45B7D1"
)

# Generate scatter plot with linear regression trend lines
penguins %>%
  drop_na(body_mass_g, flipper_length_mm) %>%
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = species, fill = species)) +
  geom_point(size = 3, alpha = 0.6, shape = 21, color = "black", stroke = 0.5) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.15, size = 1.2) +
  scale_color_manual(values = species_colors) +
  scale_fill_manual(values = species_colors) +
  labs(
    title = "Body Mass vs Flipper Length by Species",
    subtitle = "Palmer Penguins Dataset with Linear Regression Trends",
    x = "Flipper Length (mm)",
    y = "Body Mass (g)",
    color = "Species",
    fill = "Species"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "gray60"),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(face = "bold")
  )
```

```{r}
# Optional: Save the plot
ggsave("flipper_mass_regression.png", width = 10, height = 7, dpi = 300)
```

